---
title: "Knecht Vignette"
author: "Benjamin W. Campbell"
date: "`r Sys.Date()`"
geometry: margin=2cm
output: rmarkdown::html_vignette
bibliography: egoTERGM.bib
vignette: >
  %\VignetteIndexEntry{Knecht Vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Vignette Info and Introduction
  This vignette uses the `Knecht` data from Knecht (2006) and Knecht (2008) to provide an illustrative example of estimating an ego-Temporal Exponential Random Graph Model (ego-TERGM) using the `egoTERGM` package [@knecht2006networks, @knecht2008friendship].  It begins by estimating an ego-TERGM on the Knecht friendship network, a network of 26 students in a Dutch classroom measured across four time-periods.  In addition to data on the friendships between students there is nodal attribute data on a variety of demographic and behavioral factors.  For a detailed explanation of the data, I refer the reader to the original publications on the data or to the `xergm.common` help page for the `knecht` data: `??knecht`.  

Much of the code for estimating the ego-TERGM is a modification of code originally produced for its cross-sectional equivalent [@salter2015role]. This vignette will proceed step-by-step in presenting the typical workflow for estimating and interpreting an ego-TERGM using `egoTERGM`.  

*Note: The code presented here is not built given the runtime for the ego-TERGM presented here.  As such, the code presented is purely illustrative.  It will successfully run, however.* 
  
## Importing Knecht Network
  
  The following chunk of code imports the Knecht friendship network gathered by Andrea Knecht and provided in the `xergm.common` package [@leifeld2014xergm, leifeld2017temporal].  It also uses some code from the `xergm.common` package to prepare the network for the `ego_tergm` function, note that it produces a list of network objects.

```{r, eval = FALSE}
library(xergm.common)

set.seed(1)

data("knecht")

for (i in 1:length(friendship)) {
  rownames(friendship[[i]]) <- paste("Student.", 1:nrow(friendship[[i]]), sep="")
  colnames(friendship[[i]]) <- paste("Student.", 1:nrow(friendship[[i]]), sep="")
}
rownames(primary) <- rownames(friendship[[1]])
colnames(primary) <- colnames(friendship[[1]])
sex <- demographics$sex
names(sex) <- rownames(friendship[[1]])

# step 2: imputation of NAs and removal of absent nodes:
friendship <- xergm.common::handleMissings(friendship, na = 10, method = "remove")
friendship <- xergm.common::handleMissings(friendship, na = NA, method = "fillmode")

# step 3: add nodal covariates to the networks
for (i in 1:length(friendship)) {
  s <- xergm.common::adjust(sex, friendship[[i]])
  friendship[[i]] <- network::network(friendship[[i]])
  friendship[[i]] <- network::set.vertex.attribute(friendship[[i]], "sex", s)
  idegsqrt <- sqrt(sna::degree(friendship[[i]], cmode = "indegree"))
  friendship[[i]] <- network::set.vertex.attribute(friendship[[i]],
                                                   "idegsqrt", idegsqrt)
  odegsqrt <- sqrt(sna::degree(friendship[[i]], cmode = "outdegree"))
  friendship[[i]] <- network::set.vertex.attribute(friendship[[i]],
                                                   "odegsqrt", odegsqrt)
}
sapply(friendship, network::network.size)

net <- friendship
rm(list=setdiff(ls(), "net"))

```

## Estimation of ego-TERGM
Once the data is imported and transformed into a `list` of `network` objects with relevant attributes appended to each network in that list, we can proceed with estimating the ego-TERGM using `ego_tergm`.  Underneath the hood, quite a bit is occuring when the user calls this function.  First, the function transforms the list of longitudinally observed networks into a list of longitudinally observed ego-networks.  From that point, the initialization proceedure described by Campbell (2018) is used, including the estimation of a TERGM via bootstrapped pseudolikelihood and subsequent clustering of TERGM parameters using k-means clustering [@campbell2018detecting].  Upon initialization, an Expectation Maximization (EM) algorithm is used to find group-centroids and role assignments.  

The `ego_tergm` function requires a few arguments:

* `net`: The list of network objects previously discussed.
* `form`: A vector of comma-seperated `ergm-terms` used to distinguish between role assignments.
* `core_size`: The number of alters out from the ego to include and connections among them.
* `min_size`: The minimum size that an ego-network at a particular time period must achieve to be included.
* `roles`: The number of roles to be fit.
* `add_drop`: Whether nodes drop in or drop out of the network.
* `directed`: If the network is directed or not.
* `edge_covariates`: In the `form` call, are edge or dyad covariates required?
* `seed`: Seed set for replication.
* `R`: The number of bootstrap replications for the TERGM to be used to generate initial values.
* `forking`: Should parallel computing via forking be used? 
* `ncpus`: If parallel computing is to be used, how many CPUs should be used.
* `steps`: The maximum number of steps that the EM process should go through.
* `tol`: The acceptance level to assess the convergence of the EM algorithm.

```{r, eval = FALSE}
library(egoTERGM)
ego_tergm_fit <- ego_tergm(net = net,
                           form = c("edges", "mutual", "triangle", "nodeicov('idegsqrt')", "nodeocov('odegsqrt')",  "nodematch('sex')"),
                           core_size = 1,
                           min_size = 5,
                           roles = 3,
                           add_drop = TRUE,
                           directed = TRUE,
                           edge_covariates = FALSE,
                           seed = 12345,
                           R = 10,
                           forking = FALSE,
                           ncpus = 1,
                           steps = 50,
                           tol = 1e-06)

```

## Interpreting ego-TERGM Output
Once the ego-TERGM is fit using the `ego_tergm` function, there are two auxilliary functions that can be used to assist in interpreting model results.  First, I have provided an `interpret_ego_tegm` function that cleans up the group-based centroids estimated.  A warning message is coded into the function that emphasizes that these centroids should not be interpreted as TERGM parameters, but instead, as a means of understanding the distinctions between role assignments.  Second, there is a `plot_ego_tergm` function that plots the role assignments for the reduced networks.  The code is as follows.
```{r, eval = FALSE}
interpret_ego_tergm(ego_tergm_fit = ego_tergm_fit)

plots <- plot_ego_tergm(ego_tergm_fit = ego_tergm_fit)
plots[[1]]

```

## Assessing Role Generative Structure
To assist in interpreting the role generative structure of role assignments, I have included a simple function, `prepare_for_tergm`, that takes the output of an `ego_tergm` call and transforms it into a format that can be used by the `btergm` function [@leifeld2017temporal].  The code for this is as follows:
```{r, eval = FALSE}
net_list <- prepare_for_tergm(ego_tergm_fit = ego_tergm_fit)

# Indexing of the output for prepare_for_tergm refers to the role numbering from initial ego_tergm_fit
role1_btergm <- btergm(net_list[[1]] ~ edges + mutual + triangle + nodeicov('idegsqrt') +
                                        nodeocov('odegsqrt') + nodematch('sex'),
                        R = 500)

```


# References

